### 2015-06-01 WTC3/WTC3/index.Rmd
removed lock and added folders and files

---
title       : Are tree canopies optimized for carbon gain?
subtitle    : How wasteful water use in shade leaves constrain theoretical relationships of photosynthesis and resource distribution in Eucalyptus trees.
author      : Courtney Campany
job         : 
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : standalone # {standalone, draft}
knit        : slidify::knit2slides

---

##slide1

---

##slide2