### 2015-06-01 presentation/wtcIII_talk.Rmd

---
title: "Are tree canopies optimized for carbon gain?"
subtitle: How wasteful water use in shade leaves constrain theoretical relationships of photosynthesis and resource distribution in Eucalyptus trees.
author: Courtney Campany, Mark Tjoelker and Remko Duursma
date: ''
output: ioslides_presentation
widescreen: yes
smaller: yes
transition: 0.4
---

```{r knitr_options , include=FALSE}
knitr::opts_chunk$set(fig.width=6, fig.height=6,
echo=FALSE, message=FALSE,
fig.align='center', warning=FALSE)
```

```{r, results='asis'}
library(knitr)
opts_knit$set(root.dir = '../')
```

```{r global data/sourcing , include=FALSE}
source("functions and packages/packages_md.R")
source("functions and packages/functions.R")
source("functions and packages/plot_objects_all.R")

gmes <- read.csv("calculated_data/gmes_WTC.csv")
gmpair<- read.csv("calculated_data/gmes_wtc_pair.csv")
gmwet <- read.csv("calculated_data/gmes_wellwatered.csv")
Ci_bar <- read.csv("calculated_data/Ci_bar.csv")
 ###for analysis first subset well watered and drought treatments
 gm_drought <- gmes[gmes$drydown == "drought",]
 gm_water <- gmes[gmes$drydown == "control",]

leaftraits <- read.csv("calculated_data/leaftraits_summary.csv") 
par <- read.csv("raw data/par.csv")
treatments <- read.csv("raw data/chamber_trt.csv")

aci_leaf <- read.csv("calculated_data/tdl_aci.csv")
names(aci_leaf) <-c("Leaf", "Temperature", "Vcmax", "Jmax", "Vcmax_se", "Jmax_se")

  #simulate ACi curves for each leaf, ambient and elevated T
  sunAT_sim <- Aci(Ci=seq(50,1000,length=101), Vcmax=aci_leaf[3,3], Jmax=aci_leaf[3,4],PPFD=1408)
  sunET_sim <- Aci(Ci=seq(50,1000,length=101), Vcmax=aci_leaf[4,3], Jmax=aci_leaf[4,4], PPFD=1408)
  shaAT_sim <- Aci(Ci=seq(50,1000,length=101), Vcmax=aci_leaf[1,3], Jmax=aci_leaf[1,4], PPFD=375)
  shaET_sim <- Aci(Ci=seq(50,1000,length=101), Vcmax=aci_leaf[2,3], Jmax=aci_leaf[2,4],PPFD=375)
```

---
## When scaling to the canopy...Don't LEAF things to chance

- Trees can't put all leaves in the sun
- Drives changes in anatomy, morphology and physiology
- results in ↓ photosynthetic capacity in shade leaves

![](leaf_light.jpg)

- This should result in a gradient in resources available for photosynthetic with light....in theory

---

```{r, aci plot}
#plot
plot(sunAT_sim$Ci, sunAT_sim$ALEAF, col="forestgreen", pch=21,  cex=1.1,xlab=cilab, ylab="", type="l", lwd=2, ylim=c(0,30))
  points(sunET_sim$Ci, sunET_sim$ALEAF, col="forestgreen", pch=21,  cex=1.1,xlab="Ci", ylab="", type="l", lwd=2, lty=2)
  points(shaAT_sim$Ci, shaAT_sim$ALEAF, col="yellow4", pch=21,  cex=1.1,xlab="Ci", ylab="", type="l", lwd=2)
  points(shaET_sim$Ci, shaET_sim$ALEAF, col="yellow4", pch=21,  cex=1.1,xlab="Ci", ylab="", type="l", lwd=2, lty=2)
  legend("bottomright", leglab2, lty=c(1,2,1,2), lwd=2,col=colaci, pt.bg=col4,inset = 0.03)
  title(ylab=satlab, mgp=ypos)
```
---