### 2015-06-11 markdown/wtc_gm_compare.Rmd
re ran gmes with tobacco parameters
created new gmes calc function with tob

have new RMD that just compares those values

tweaked c13 fig

---
title: "Sun vs Shade leaves (WTCIII)"
author: "Court Campany"
date: "April 2, 2015"
output: html_document
---
```{r knitr_options , include=FALSE}
knitr::opts_chunk$set(fig.width=6, fig.height=6,
echo=FALSE, message=FALSE,
fig.align='center', warning=FALSE)
```

```{r, results='asis'}
library(knitr)
opts_knit$set(root.dir = '../')
```

```{r global data/sourcing , include=FALSE}
source("functions and packages/packages_md.R")
source("functions and packages/functions.R")
source("functions and packages/plot_objects_all.R")

#gm calc from E.globulus
gmes <- read.csv("calculated_data/gmes_WTC.csv")
gmpair<- read.csv("calculated_data/gmes_wtc_pair.csv")
###for analysis first subset well watered and drought treatments
 gm_drought <- gmes[gmes$drydown == "drought",]
 gm_water <- gmes[gmes$drydown == "control",]

#gm calc from tobacco 
gmes_tob <- read.csv("calculated_data/gmes_WTC_tobacco.csv")
gmpair_tob<- read.csv("calculated_data/gmes_wtc_tobacco_pair.csv")

###for analysis first subset well watered and drought treatments
 gm_drought_tob <- gmes_tob[gmes_tob$drydown == "drought",]
 gm_water_tob <- gmes_tob[gmes_tob$drydown == "control",]

```

#### <u>Mesophyll Conducatnce</u> from Euc parameters

```{r,results='asis'}
###make a table of means and se of gm

 gm_agg <- summaryBy(gm~ leaf+light+temp+leaflight, data=gm_water,FUN=c(mean,se))
   tablenames <- c( "Leaf Type", "Light", "Temperature","leaflight","Gm", "Gm(se)")
   names(gm_agg) <- tablenames
  gm_agg2 <- gm_agg[c(3,4,5,6,1,2),]
  row.names(gm_agg2) <- NULL

panderOptions("digits", 3)
pander(gm_agg2,caption="Well Watered")

  gm_dr <- summaryBy(gm~ leaf+light+temp+leaflight, data=gm_drought,FUN=c(mean,se))
    tablenames <- c( "Leaf Type", "Light", "Temperature","Leaf-Light","Gm", "Gm(se)")
    names(gm_dr) <- tablenames
  gm_dr2 <- gm_dr[c(3,4,5,6,1,2),]
  row.names(gm_dr2) <- NULL

pander(gm_dr2,caption="Drought")
```

#### <u>Mesophyll Conducatnce</u> from tobacco parameters

```{r,results='asis'}
###make a table of means and se of gm

 gm_agg_tob <- summaryBy(gm~ leaf+light+temp+leaflight, data=gm_water_tob,FUN=c(mean,se))
   tablenames <- c( "Leaf Type", "Light", "Temperature","leaflight","Gm", "Gm(se)")
   names(gm_agg_tob) <- tablenames
  gm_agg3 <- gm_agg_tob[c(3,4,5,6,1,2),]
  row.names(gm_agg3) <- NULL

panderOptions("digits", 3)
pander(gm_agg3,caption="Well Watered")

  gm_dr_tob <- summaryBy(gm~ leaf+light+temp+leaflight, data=gm_drought_tob,FUN=c(mean,se))
    tablenames <- c( "Leaf Type", "Light", "Temperature","Leaf-Light","Gm", "Gm(se)")
    names(gm_dr) <- tablenames
  gm_dr3 <- gm_dr_tob[c(3,4,5,6,1,2),]
  row.names(gm_dr3) <- NULL

pander(gm_dr3,caption="Drought")
```