### 2015-07-01 markdown/WTC3_figures.Rmd
plot changes

---
title: "WTCIII-Figures"
author: "Court Campany"
date: "June 23, 2015"
output: html_document
---

```{r knitr_options , include=FALSE}
knitr::opts_chunk$set(fig.width=6, fig.height=6,
echo=FALSE, message=FALSE,
fig.align='center', warning=FALSE)
```

```{r, results='asis'}
library(knitr)
opts_knit$set(root.dir = '../')
```

```{r global data/sourcing , include=FALSE}
source("functions and packages/functions.R")
source("functions and packages/packages.R")
source("master_scripts/plot_objects.R")
library(pander)

#treatments <- read.csv("raw data/chamber_trt.csv")
```

#### <u>Design</u>
* WTCIII was a warming experiment with ambient and +3°C warming (n=6).  A drought manipulation was added in the last few months of the experiment but is not included here.
* Sun leaves were measured in the upper canopy at their current PPFD.  Shade leaves were measured in the lower/middle canopy, first at their current PPFD and then again at sun leaf PPFD as a simulated sunfleck.
* New fully expanded leaves were measured. Shade leaves were sampled into the middle canopy as the experiment progressed to avoid issues with leaf age.
* Leaf physiological measurements were taken at their current temperature (ambient, +3°C). Thus, data below represent a large range of temperatures from 18-35°C.

#### <u>Figures</u>
```{r par, fig.width=6, fig.height=6}
source("master_scripts/par.R")
```  
Figure 1. Local light environment for sun and shade leaves for each gas exchange campaign. Means and standard errors represent 1m integrated PPFD (n=5), measured with a ceptometer, at the canopy height of each selected leaf.

```{r acicurves, fig.width=8, fig.height=10}
source("master_scripts/aci_vcmaxnitro.R")
```  
Figure2. Predicted photosynthetic CO~2~ response (ACi) curves for sun and shade leaves at elevated and ambient temperature treatments. ACi curves were developed at 25°C and at saturating light (1800 $\mu$mols m^-1^ s^-1^)(a). The relationship between Vc~max~ and leaf nitrogen~area~ for sun and shade leaves at elevated and ambient temperature treatments (b). ACi curves were taken once on well watered trees before the initiation of the drought treatment.

```{r wateruse, fig.width=8, fig.height=10}
source("master_scripts/wateruse.R")
```  
Figure 3. Response of instantaneous water use efficiency (ITE) to leaf vapor pressure deficit (VPD) for sun leaves and shade leaves at both low and high light (a) and the relationship between leaf $\delta$^13^C and leaf nitrogen~area~ for sun and shade leaves at elevated and ambient temperature treatments. For (a) VPD is the leaf to air pressure difference inside the gas exchange cuvette and lines represent predictions from the optimal ITE model with a g1 value for each leaf type and treatment.

```{r gmgs, fig.width=8, fig.height=10}
source("master_scripts/gmgs.R")
```  
Figure 4. The response of leaf photosynthesis rates to stomatal conductance (a) and mesophyll conductance (b) for sun leaves and shade leaves at both low and high light with elevated and ambient temperature treatments.  Dashed lines represent 95% confidence intervals from a general additive model with a smoothing factor of 5.  The response of shade leaf physiology to high light was recorded once both photosynthesis and stomatal conductance were stable in the gas exchange leaf cuvette (ca. 25min).

```{r cicc, fig.width=10, fig.height=6}
source("master_scripts/cicc_bar.R")
```  
Figure 5. The average intercellular CO~2~ concentraion (a), CO~2~ concentration in the chloroplasts (b) and CO~2~ drawdown from substomatal cavities to sites of carboxylation of sun leaves and shade leaves at both low and hight light with standard errors.

```{r barplots, fig.width=10, fig.height=6}
source("master_scripts/physiology_barplots.R")
```  
The average stomatal conductance (a), mesophyll conductance (b) and photosynthesis (c) between sun leavse and  shade leaves at both low and high light with standard errors.

#### <u>Supplementary Figures</u>
```{r gmparN, fig.width=8, fig.height=10}
source("master_scripts/gm_par_nitro.R")
``` 
Figure S1. The relationship between mesophyll conductance, leaf PPFD and leaf nitrogen.

```{r gmT, fig.width=6, fig.height=6}
source("master_scripts/gm_temp.R")
``` 
Figure S2. Response of mesophyll conductance to leaf temperature  for sun and shade leaves at elevated and ambient temperature treatments. For (a) and (b) lines represent 95% confidence intervals generated with bootstrapping methods (n=999) from mixed effect linear model equations with chamber as a random effect.