### 2015-12-01 manuscript/manuscript_wtc3.Rmd
abbreviation updates to manuscript

---
title: "Rapid response of mesophyll conductance to light availability allows shade leaves to take advantage of sunflecks"
author: ""
date: ""
output:
  word_document:
    reference_docx: manuscriptstyle.docx
  pdf_document: default
  html_document:
    number_sections: yes
csl: new-phytologist.csl
bibliography: references.bib
---

\
Courtney E. Campany^1^, Belinda Medlyn^1^, Mark Tjoelker^1^, Susanne von Caemmerer^2^, Remko A. Duursma^1^.

\
^1^ Hawkesbury Institute for the Environment, Western Sydney University, Locked Bag 1797, Penrith, NSW, Australia

\
^2^ Plant Science Division, Research School of Biology, The Australian National University, Canberra, ACT, Australia

\
*Corresponding author*: 
Courtney Campany
E: courtneycampany@gmail.com

```{r, results='asis', echo=FALSE, warning=FALSE}
library(knitr)
opts_knit$set(root.dir = '../')
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Two useful packages for printing numbers, accessing content of objects.
library(broom)
library(reporttools)

# Set Flag to include figures in text or not.
includeFigs <- TRUE

# This document can be compiled with the 'Knit Word' button in Rstudio, or:
# library(rmarkdown)
# render("manuscript.Rmd", "word_document", "manuscript.docx")
```

```{r global data/sourcing , include=FALSE}
source("functions and packages/functions.R")
source("functions and packages/packages.R")
source("master_scripts/plot_objects.R")
library(pixiedust)

##read table ready dataframes
resourcetab <- read.csv("master_scripts/resource_table.csv", stringsAsFactors = FALSE)
getab <- read.csv("master_scripts/ge_table.csv", stringsAsFactors = FALSE)
pval <- c(0.785, 0.001, 0.028, 0.002, 0.001, 0.3486, 0.6385, 0.973)
pval2 <- c(0.001,	0.001,	0.001,	0.001,	0.001,	0.005,	0.001,	0.001)
# getab <- read.csv("master_scripts/ge_table.csv")

#treatments <- read.csv("raw data/chamber_trt.csv")
```

```{r, echo=FALSE}

parfig <- "1"
acifig <- "2"
agmgsfig <- "3"
gebarfig <- "4"
waterusefig <- "5"
deltafig <- "6"
ciccfig <- "7"

airvarsfig <- "S1"
accfig <- "S2"
get1fig <- "S3"
get2fig <- "S4"

```

# Summary
* Light gradients within tree canopies play a major role in the distribution of plant resources that define photosynthetic capacity of individual leaves. A lack of empirical data relating photosynthesis to leaf physiological behavior within tree canopies, however, impedes our ability to assess the contribution of shade leaves to canopy carbon gain.
* To investigate the CO~2~ diffusion pathway of sun and shade leaves, leaf gas exchange was coupled with online carbon isotope discrimination to measure photosynthesis (*A*), stomatal conductance (g~s~) and mesophyll conductance (g~m~) in *Eucalyptus tereticornis* trees grown in climate controlled whole tree chambers.
* Compared to sun leaves, shade leaves had lower A, g~m~, Vc~max~, J~max~ and leaf nitrogen but maintained similar rates of g~s~. When light intensity was increased from low light to high light for shade leaves both g~s~ and g~m~ increased rapidly, leading to increases in A greater than sun leaves at the same PPFD. 
* Here we show that dynamic physiological responses of shade leaves to altered light environments has important implications for upscaling leaf level measurements and predicting whole canopy carbon gain. We argue that the rapid response of g~m~ with light enables shade leaves to respond quickly to sunflecks. Evidence that g~m~ not only varies within a canopy but can be up-regulated over short time intervals possibly represents a new mechanism underpinning leaf gas exchange responses to light.

## Key words
photosynthesis, stomatal conductance, mesophyll conductance,  shade, leaf optimal behavior

# Introduction
Light availability is one of the most important environmental drivers of leaf carbon (C) uptake in trees. Predicting C uptake if forests usually involves upscaling leaf level measurements to assess whole canopy function. Due to the costs and limitations of efficient light harvesting, plants cannot expose all leaves to full sun [@niinemets2010review], making simple upscaling based on solar irradiance problematic. Incident photosynthetic photon flux density (PPFD) declines exponentially with cumulative leaf area index, creating a steep light gradient from the canopy top to bottom [@monsi2005factor]. Consequently, structural and functional properties of leaves are modified to efficiently use light [@vogelman1996leaves; @valladares2004photosynthetic], as changing irradiance strongly affects rates of photosynthesis (*A*) [@evans1995carbon]. To estimate whole canopy C gain it is thus necessary to account for the non-linear response of A to light by distinguishing between shaded and sunlit leaves [@de1997simple; @linderson2012up].

\
The distribution of resources required for *A*, including leaf nitrogen (N) and supply of water, are also partially defined by canopy light gradients. As *A* has a saturating response with light and maximum rates depend in part on N-rich photosynthetic machinery, it has been argued that leaf N should be proportional to PPFD along the canopy light gradient to maximize canopy C gain at a given total canopy N [@field1983allocating; @field1986photosynthesis; @peltoniemi2012co]. Changes in chlorophyll per unit N, chlorophyll a:b ratios, electron transport capacity per unit chlorophyll, and ratios of electron transport capacity to Rubisco activity can also occur in response to changes in irradiance [@evans2001photosynthetic]. Sun leaves frequently experience greater water limitations in the upper canopy, despite effective vascular systems developed for high radiation loads and transpiration [@sellin2008effects; @niinemets2012optimization]. Higher rates of *A* and stomatal conductance (g~s~) can only be sustained if the hydraulic conductance (K) is also large enough to avoid low leaf water potentials [@hubbard2001stomatal]. Optimal photosynthetic N investment in the upper canopy will be ineffective in enhancing *A* if water supply is insufficient [@niinemets2012optimization; @peltoniemi2012co], thus K should also be higher in the upper canopy to supply sunlit leaves with sufficient water [@burgess2006hydraulic;@sellin2007effects ;@sellin2008effects].

\
Rates of photosynthesis in C~3~ plants are limited by the [CO~2~] available for fixation by Rubisco within the chloroplast and this [CO~2~] is a function of the drawdown of CO~2~ from the atmosphere to the site of carboxylation [@warren2008stand]. This drawdown consists of multiple resistance pathways to CO~2~ diffusion which include CO~2~ diffusion from the atmosphere through stomata (stomatal conductance, g~s~) and then from these intercellular air spaces into the chloroplast (mesophyll conductance, g~m~). Based on optimal theory, regulation of g~s~ within a tree canopy should act to efficiently utilize available supplies of light, N and water to maximize *A* [@peltoniemi2012co]. This is because stomata are hypothesized to exhibit an optimal behaviour to maximize C gain while simultaneously minimizing water loss through transpiration [@cowan1977stomatal]. Mesophyll conductance can also impose limitations on *A* as large g~s~ [@warren2008stand; @ubierna2011estimation], reducing the efficiency of leaf N use in *A* [@niinemets2007photosynthesis] if g~m~ constrains CO~2~ supply to the chloroplast. Part of the variation in photosynthetic capacity between sun and shade leaves has been proposed to be due to differences in g~m~ [@piel2002effect; @warren2007internal], yet the trade-offs that constrain this diffusion pathway are yet to be explicitly quantified. Stomatal and mesophyll conductance should not be considered independent of each other [@griffiths2013mesophyll], but a lack empirical data currently hinders are ability to interpreting their coupled responses to *A* across sun and shade leaves.

\
Additionally, assessing shade leaf behaviour is made difficult with accounting of short term light fluctuations within a canopy, via sunflecks. How shade leaves utilize sunflecks for short term C gain depends on the combined response time of g~s~ and g~m~ and the underlying photosynthetic biochemistry acclimated to a low light environment [@pearcy1990sunflecks;@tausz2005dynamic]. The utilization of sunflecks is first limited by delayed responses of stomata opening, which may take minutes, effectively limiting the maximum assimilation rate that can be achieved [@pearcy1990sunflecks; @vico2011effects; @way2012sunflecks]. Mesophyll conductance has been shown to respond to environmental factors (e.g. CO~2~, temperature or vapor pressure deficit) at timescales of minutes, possibly faster than g~s~ [@flexas2008mesophyll and references therein], yet short term response to light availability are unclear. For example, g~m~ was found to be independent of light intensity in wheat leaves [@tazoe2009light] but was responsive to light in tobacco [@flexas2007rapid]. Anatomical parameters which regulate g~m~ with changing irradiance such as chloroplast surface area [@terashima2006irradiance] and mesophyll thickness [@boardman1977comparative] are also unlikely to adjust during short light fluctuations. The physiological behaviour of shade leaves to maximize C gain must be assessed as both a degree of acclimation to local irradiance and as a potential response to transitory light availability.

\
Climate warming may also affect the physiological behaviour of leaves within a canopy. This is because leaves can be exposed to different heat, water and high light stresses as temperature and vapour pressure deficit (VPD) vary with canopy light availability [@baldocchi2002environment; @valladares2004photosynthetic; @niinemets2007photosynthesis]. How these stresses affect the diffusion of CO~2~, through either g~s~ or g~m~, will have implications for upscaling *A* for sun and shade leaves. Additionally, light saturated rates of A are limited by the maximum rate of Rubisco carboxylation (Vc~max~) or the maximum rate of photosynthetic electron transport (J~max~) across a range of temperatures, yet their temperature dependencies are not the same [@farquhar1980biochemical; @medlyn2002temperature]. How these parameters are differentially affected by warming may impact constraints of N distribution and leaf photosynthetic capacity across light gradients. The impacts of warming on plant physiological processes are obviously vast, yet differentiating their impacts on leaf physiology within a canopy will be essential to evaluate whole tree responses to a changing climate.

\
In this study we use *Eucalyptus tereticornis* Sm. trees, planted in climate controlled whole tree chambers with ambient and elevated temperature (ambient +3&deg;C) treatments, to empirically evaluate the distribution of nitrogen and water supply and leaf physiological behaviour of sun and shade leaves. Our hypotheses are as follows:

\
1.	If whole tree canopies are optimized for C gain, then leaf N, hydraulic conductance and biochemical photosynthetic capacity are predicted to be higher in sun leaves compared to shade leaves.

\
2.	Stomatal conductance is proportional to *A* across sun and shade leaves under similar leaf VPD and g~m~ scales positively with photosynthetic capacity.

\
3. As shade leaves were expected to develop lower biochemical photosynthetic capacity and leaf physiological responses to light are typically slow (minutes), increases in *A* following sunfleck simulations were not expected to reach rates of full sun leaves.

\
4.	The effects of climate warming were predicted to be greater in sun than shade leaves, seen as a decrease in g~s~ and leaf C gain during summer months, as increased evaporation demand from higher temperatures will lead to stomatal closure.

# Materials and Methods

## Whole tree chamber experimental design
Twelve *Eucalyptus tereticornis* seedlings, chosen from a single local Cumberland plain cohort, were planted in March 2013 into 12 whole-tree chambers (WTC) at the Hawkesbury Forest Experiment site near Richmond, NSW, Australia. Each chamber has a height of 9 m and seedlings were grown for 15 months. A detailed description of the WTC operation and design is available in [@barton2010whole]. Six chambers were set to match outside ambient air temperatures (AT) while the remaining 6 experienced an elevated air temperature treatment of +3&deg;C (ET, Figure `r airvarsfig`). Trees grew quickly and developed large canopies, with height growth reaching the top of the WTCs over the experiment duration. Trees were watered weekly with 70 L from March 2013 to November 2013. From October 2013 to the end of the experiment trees were watered every 15 days with the mean monthly rainfall amount for Richmond, NSW. In February 2013 half of the chambers (3 each of AT and ET) were subjected to a drought treatment by withholding watering. Due to a limited range of data for the drought treatment only well-watered trees are reported, which reduces the sample size of WTC (n=6 to n=3) for the final 3 months of the experiment.

\
Before seedlings were planted into each chamber they were maintained under well-watered conditions in 25 L pots and kept inside each chamber. This allowed for seedlings to gain sufficient size before planting while also allowing them to acclimate to chamber temperature treatments. Seedlings were planted into each chamber after mean seedling height reached 100 cm. The top soils at this site, used in both pots and chambers, are an alluvial formation of low-fertility sandy loam soils (380 and 108 mg kg-1 total N and phosphorus respectively) with low organic matter (0.7 %) and low water holding capacity. A root exclusion barrier extended from chamber walls to the hard layer (ca. 1 m) and roots were allowed to grow freely below the barrier. Leaf gas exchange measurements were initiated in October 2013 when trees had both ample height growth and canopy development for realistic canopy light gradients to be measured. At this point, trees under AT treatment had a mean diameter of 28.2&pm;1.1 mm, height of 348&pm;15.1 cm and an leaf area of 3.9&pm;0.1 m^2^. For ET treatments, trees had a mean diameter of 34.1&pm;2.1 mm, height of 418.3&pm;23.1 cm and an leaf area of 6.2&pm;0.2 m^2^. Leaf area was calculated based on leaf counts and mean leaf size.

##Leaf gas exchange, online carbon isotope discrimination and mesophyll conductance
Leaf gas exchange measurements were performed six times, beginning in October 2013 and monthly from December 2013 to April 2014. Measurements were taken on a representative sun and shade leaf for each tree during each measurement campaign. The newest fully expanded leaf from the branch apex was chosen for gas exchange measurements and sun leaves were measured in the upper third of the canopy. Here, shade leaves are defined as inner-canopy leaves developing on secondary branches in a low light environment. Shade leaves were always measured in the lower canopy, but leaves were sampled on subsequent higher branches across measurement campaigns to minimize confounding effects of leaf age. As shade leaves most likely developed slower this assured that older leaves in the bottom canopy were avoided.

\
Prior to gas exchange measurements photosynthetic photon flux density (PPFD) was recorded as a point measurement at the individual leaf level and a spatially averaged measurement at the canopy position for each selected leaf.  A hand-held quantum sensor (LI-COR, Lincoln, NE, USA) was used to record leaf level PPFD to ensure that chosen leaves were positioned in the desired light environment, either full sun or full shade. A ceptometer (AccuPAR LP-80, Decagon Devices, Pullman, USA) was then used to measure a spatially averaged PPFD at the canopy height of each chosen leaf type. Each ceptometer reading integrated an array of 80 sensors over a total length of 84 cm. Five ceptometer readings were recorded at different locations within the canopy, but at the same height and close to each selected leaf. The mean of these readings was assumed to represent the local light environment of sun and shade leaves for each tree. All measurements of PPFD and gas exchange were performed on sunny days between 10:00-14:30 h.

\
Leaf level gas exchange was measured with a standard (2 x 3 cm) leaf cuvette using a portable gas exchange system (LI-6400XT, LI-COR, Lincoln, NE, USA).  This system was coupled with a tunable diode laser (TDL; TGA100,Campbell Scientific, Inc., Logan, UT, USA) for concurrent measurements of online C isotope discrimination. The CO~2~ in the leaf cuvette was set at ambient atmospheric [CO~2~] (400 ppm) with a flow rate of 200 $\mu$mol s^-1^. Two identical gas exchanges systems were run simultaneously, one in each of a randomly chosen WTC for each temperature treatment. Leaf temperatures were controlled at the current AT or ET WTC air temperature. PPFD in the cuvette was set to match the individual light environment of each leaf type (explained above). Periods of high irradiance were simulated for shade leaves by increasing the leaf cuvette PPFD (LI-COR red/blue light source) to match the light environment of the full sun leaf in the same tree. The maximum sunfleck response of shade leaves was then recorded once CO~2~ and water vapor fluxes re-stabilized in the leaf cuvette (ca. 25 min).

\
Once CO~2~ and water vapor flux values were stable for each leaf measurement, the sample and reference gas lines were diverted to the TDL via T-junctions inserted into the reference gas tube and match valve outlet of the LI-6400. The gas streams were dried by passing through napion gas dryers in the respective gas lines, and then ^12^CO~2~ and ^13^CO~2~ concentrations were measured for each gas stream by the TDL. Reference, sample and two calibration gases were run on alternating 80 s loops (20 s each), one for each AT and ET leaf at a matched canopy position, for a total of 12 min. This allowed for 4-5 measurements per leaf and data were averaged over the last 10 s of reference line and samples line gas streams for calculations. The two calibration gases were drawn from compressed air tanks (330 and 740 ppm CO~2~) in order to correct for gain drift of the TDL on each measurement cycle. Photosynthesis, g~s~, transpiration, VPD and intercellular [CO~2~] (C~i~) were auto-logged every 15 s for each gas exchange system over the same 12 min interval.

\
Using online C isotope discrimination measurements, the difference  between the observed discrimination and what is predicted for light saturated gas exchange is proportional to g~m~ [@griffiths2013mesophyll]. First, leaf C isotope discrimination was calculated by comparing the isotopic composition of the reference gas entering the leaf cuvette ($\delta$^13^C~e~) with the sample gas ($\delta$^13^C~o~) such that:

\
(1)
$$\delta = \left(\frac{R_s}{R_{stnd}} -1\right)1000$$

where R~s~ is the isotopic ratio of the sample and R~stnd~ is the isotopic ratio of the standard Vienna Pee Dee Belemnite (VPDP). Next, the observed discrimination ($\Delta$~obs~) is calculated from Evans et al. [-@evans1986carbon]:

\
(2)
$$\Delta_{obs} = \frac{\xi(\delta^13C_o-\delta^13C_e)}{1000+\delta^13C_e-\xi(\delta^13C_o-\delta^13C_e)}$$

where:

\
(3)
$$\xi = \frac{C_{e}}{C_{e}-C_{o}}$$

and $\xi$ is the ratio of the CO~2~ entering the well mixed leaf cuvette to the CO~2~ draw down in the gas stream by the leaf.

\
Second, C isotope discrimination during C~3~ photosynthesis ($\Delta$) is the resultant discrimination from CO~2~ diffusion from the atmosphere to the site of carboxylation, consisting of a series of fractionation steps described in Evans et al. [-@evans1986carbon]. In this experiment, a modified form of this equation presented in Evans & Von Caemmerer [-@evans2013temperature] with ternary effect corrections by Farquhar & Cernusak [-@farquhar2012ternary] was used such that:

\
(4)
$$\Delta_o = \Delta_i - \Delta_{gm} - \Delta_e - \Delta_f$$

where $\Delta$~o~ is the observed discrimination and  $\Delta$~i~, $\Delta$~gm~ , $\Delta$~e~ and $\Delta$~f~ are the contributions to fractionation if C~i~ = C~c~, g~m~, respiration and photorespiration, respectively. The equations for each are as follows:

\
(5)
$$
\begin{aligned}
\Delta_i = &\frac{1}{1-t}a_i+\frac{1}{1-t}((1+t)b-a')\frac{C_i}{C_a}
\end{aligned}
$$

\
(6)
$$
\begin{aligned}
\Delta_{gm} = &\frac{1+t}{1-t}\left(b-a'-\frac{eR_d}{A+R_d}\right)\frac{A}{g_mC_a}
\end{aligned}
$$

\
(7)
$$
\begin{aligned}
\Delta_e = &\frac{1+t}{1-t}\left(\frac{eR_d}{(A+R_d)C_a}(C_i-\Gamma_*)\right)
\end{aligned}
$$

\
(8)
$$
\begin{aligned}
\Delta_f = & \frac{1+t}{1-t}\left(f\frac{\Gamma_*}{C_a}\right)\\
\end{aligned}
$$

where the different fractionation factors include; diffusion through water (a~i~, 1.8&permil;), Rubisco carboxylation (b, 29&permil;), the photorespiratory fractionation (f, 16.2&permil;) and the combined fractionation through the boundary layer and the stomata (a'). a' is defined by:

\
(9)
$$a' = \frac{a_b(C_a-C_i)+a(C_s-C_i)}{C_a-C_i}$$

where C~s~ is the CO~2~ partial pressure at the leaf surface, a~b~ is the fractionation from boundary layer diffusion (2.9&permil;) and a is the fractionation due to diffusion in air (4.4&permil;) [@evans1986carbon]. C~a~ and C~i~ are the atmospheric and intercellular partial pressures and $\Gamma_{*}$ is the compensation point in the absence of mitochondrial respiration in the light (R~d~). In this experiment both $\Gamma_{*}$ and R~d~ were derived using a standard Arrhenius function with parameters for *Eucalyptus globulus* from [@crous2012light]. The ternary effect corrections (t) are described by:

\
(10)
$$t = \frac{(1+a')E}{2g^t_{ac}}$$

where E denotes the transpiration rate and $\text{g}^t_{ac}$ is the total conductance to CO~2~ diffusion to both the boundary layer and stomatal conductance.

\
The CO~2~ diffusion from the intercellular airspace to the chloroplast, g~m~, is given by its relationship to the CO~2~ assimilation rate (*A*) by:

\
(11)
$$g_m = \frac{*A*}{C_i-C_c}$$

where C~c~ is the chloroplast CO~2~ partial pressure. Once g~m~ was calculated C~c~ and the drawdown of CO~2~ from the intercellular air spaces to the site of carboxylation were then estimated using Equation 11. Examples of this approach to measure gas exchange and C isotope discrimination are presented in Evans and Von Caemmerer [-@evans2013temperature]. The variation in $\Delta$~o~ between sun and shade leaves and the simulated sunfleck where then compared as a function of C~i~:C~a~.

##Biochemical parameters of photosynthesis
Photosynthetic CO~2~ response (AC~i~) curves were measured at 25 &deg;C for one sun and shade leaf for each WTC prior to the initiation of the drought treatment. Each AC~i~ curve began at the reference [CO~2~] of 400 $\mu$l l^-1^ and then consisted of 12 additional steps from [CO~2~] of 50 to 1800 $\mu$l l^-1^ at 25 &deg;C at saturating light (1800 $\mu$mols m^-1^ s^-1^). From these curves the photosynthetic parameters, J~max~ and Vc~max~, were quantified using the biochemical model of [@farquhar1980biochemical] and fit with the 'plantecophys' package [@duursma2015] in R [@RDevelopmentCoreTeam2011].

##Leaf chemistry and hydraulic parameters
Following gas exchange measurements each leaf was collected, measured for leaf water potential (explained below), scanned for leaf area, dried and weighed. These leaves were then milled and analyzed for leaf N content and $\delta$^13^C. Leaf samples were analysed on a Delta V Advantage coupled to a Flash HT and Conflo IV (Thermo Fisher Scientific, Bremen, Germany) in dual-reactor setup. Samples were flash combusted at 1000&deg;C and converted to CO~2~ and N~2~ and then subjected to stable isotope ratio mass spectrometry. Leaf N is reported on an area basis (N~a~, g m^-2^) and isotopic signatures are reported relative to the VPDP scale.

\
Predawn ($\Psi_{pd}$) and midday ($\Psi_l$) leaf water potentials (MPa) were measured for sun and shade leaves during each gas exchange campaign using a PMS 1505D pressure chamber (PMS Instruments, Albany, OR, USA). The leaf closest to the leaf used for gas exchange was sampled for measurement of $\Psi_{pd}$. Predawn leaf water potential was measured before sunrise on the same day as gas exchange measurements. Leaves used for gas exchange were immediately sampled for $\Psi_l$ once measurements were completed. All leaves were detached and immediately stored inside foil covered bags before water potential measurements were performed. Leaf water potential and transpiration (E, mmol^-2^ m^-2^ s^-1^) from gas exchange were then used to calculate leaf-specific hydraulic conductance (K~l~, mmol m^-2^ s^-1^ MPa) through the equation:

\
(12)
$$K = \frac{E}{\Psi_{pd}~-\Psi_l}$$

Leaf level instantaneous transpiration efficiency (ITE) was calculated as A divided by E. The g~1~ parameter was estimated from ITE to VPD response curves by fitting a rearranged optimal g~s~ model for ITE [@medlyn2011reconciling] using non-linear regression, where k=0.5 [see @duursma2013near]. Small values of g~1~ indicate that transpiration is costly in C terms and plants are then likely to exhibit conservative water use [@lin2015optimal].

##Data analysis
Differences in experimental parameters to either the warming treatment or leaf type were analysed by mixed-effects models in R [@RDevelopmentCoreTeam2011] with WTC as a random effect. Explained variance (R^2^) of mixed models were computed as in  [@nakagawa2013general]. Confidence intervals (95 %) of mixed effect linear models were generated using bootstrapping methods with 999 simulations, using the bootMer function in the 'lme4' package [@bates2013lme4]. For non-linear relationships, confidence intervals were estimated by fitting a generalized additive model to the data with the 'mgcv' package, using WTC as a random effect. All tests of statistical significance were conducted at an alpha of 0.05.

# Results

## Leaf resource distribution
Across six measurement campaigns over the 7 month period PPFD was  reduced on average by >75% in the shade (Figure `r parfig`). Leaf-specific hydraulic conductance (K~l~) was similar across sun and shade leaves (Table 1). This was because $\Psi_{pd}$, $\Psi_l$ (Table 1) and E (Table 2) did not differ between leaf types. Leaf N~a~ was approximately 20% higher in sun leaves compared to shade leaves (Table1). Leaf mass per area was not different between leaf types. No effect of the warming treatment was detected with  PPFD, $\Psi_{pd}$, $\Psi_l$, K~l~, E, N~area~ or LMA either within or across leaf types.

## Photosynthetic capacity and leaf photosynthesis rates 
The photosynthetic parameters J~max~ and Vc~max~ were higher in sun compared to shade leaves (Table 1), as estimated from AC~i~ curves (Figure `r acifig`a). Within leaf types, no effect of the warming treatment was detected on either parameter. Among the sampled leaves, Vc~max~ was positively related to leaf N~a~ across leaf types and temperature treatments (P = 0.01, Figure `r acifig`b).

\
Mean *A* was significantly higher in sun compared to shade leaves (+23%), when measured at their local light environment (Table 2). Additionally, leaf N~a~ was positively related to *A* across gas exchange campaigns and leaf types measured under ambient light and temperature conditions (P < 0.001, Figure `r acifig`c). Following an increase in light intensity to match high-light conditions, A of shade leaves increased to values significantly greater than sun leaves at high light (P < 0.001, Table 2). No effect of the warming treatment was detected on *A* of sun leaves measured at high light or shade leaves at either low or high light. Photosynthesis within leaf types and treatments was similar through time and across the range of leaf temperatures measured (Figure `r get1fig`a).

## Stomatal conductance and leaf water use efficiency
On average, g~s~ was 18% higher in shade compared to sun leaves under their local light environment (Table 2). Photosynthesis was positively correlated with g~s~ in all leaves measured under high light conditions, however, g~s~ and *A* were not correlated in shade leaves under low light (Figure `r agmgsfig`a). Following increased PPFD, g~s~ of shade leaves was significantly greater than both shade leaves at low light and sun leaves, pooled across all measurement dates (Figure `r gebarfig`a). No effect of the warming treatment was detected on g~s~ within or across leaf types. Stomatal conductance within leaf types and treatments was similar through time and across the range of leaf temperatures measured (Figure `r get2fig`b).

\
Measured under ambient light and temperature conditions leaf ITE was significantly greater in sun leaves than in shade leaves at low light (+21%, P = 0.001). Following an increase in  PPFD to high-light conditions, ITE of shade leaves did not differ from shade leaves at low light and was still significantly lower than sun leaves (P < 0.001). Instantaneous transpiration efficiency in sun leaves was reduced in the warming treatment, but no warming effect was detected in shade leaves measured at low or high light (Table 2). The mean estimated g~1~ for sun leaves was 1.51&pm;0.11 and for shade leaves with low and high light was 2.59&pm;0.12 and 2.74&pm;0.04. For all leaf types and light treatments there was a strong response of ITE to VPD and individual data points broadly corresponded to response curves from the optimal ITE model with a g~1~ value for each leaf type and treatment (Figure `r waterusefig`a).  Within leaf types and light treatments the response of VPD to leaf temperature was similar across all measurement campaigns (Figure `r get2fig`a).

\
Bulk-leaf $\delta$^13^C, as an index of integrated water use efficiency [@marshall2007sources], was significantly lower in shade leaves compared to sun leaves by ca. 2&permil; (Table 1). No effects of the warming treatment on leaf $\delta$^13^C were detected. Leaf $\delta$^13^C and N~area~ were positively correlated for all leaves (P<0.001, Figure `r waterusefig`b), with less negative $\delta$^13^C (higher water-use efficiency) and higher N investment in sun leaves.

## Leaf carbon isotope discrimination and mesophyll conductance
The observed carbon isotope discrimination ($\Delta_o$) measured during photosynthesis was positively correlated with C~i~:C~a~ for all leaf types (P < 0.001), with larger $\Delta$~o~ detected for sun leaves and shade leaves at high light than shade leaves at low light (Figure `r deltafig`). Carbon isotope discrimination associated with g~m~ accounted for the majority of $\Delta_o$ (69.7&pm;0.4%) and varied little across measurement temperatures, leaf types, or warming treatments. The remainder $\Delta_o$ consists of the contributions of g~s~, respiration and photorespiration to discrimination.

\
Mean g~m~ was higher in sun compared to shade leaves (+27%) under their local light environment (P < 0.001). Following a short-term increase in PPFD from low to high light, g~m~ of shade leaves increased to values significantly greater than sun leaves (Table 2). Proportional increases in g~m~ were matched by proportional increases in *A* from low to high light in shade leaves (Figure `r gebarfig`b,c). Photosynthesis scaled positively with increases in g~m~ for all leaves, with similar intercepts but different slopes between leaf type and light treatment (P = 0.0186). The large increases in g~m~ in shade leaves under high light resulted in the highest rates of *A* (Figure `r agmgsfig`b). No differences in g~m~ were detected with the warming treatment within leaf types. Mesophyll conductance did not vary across measurements campaigns within leaf types and light treatments (Figure `r get1fig`b), but a weak negative relationship with increasing leaf temperature was detected with sun and shade leaves under their local light environment (P = 0.001 & 0.04, respectively). We also simulated AC~c~ curves to determine if treatment differences in J~max~ and Vc~max~ where instead the result of differences in g~m~. Comparison of A-Cc curves (Figure `r accfig`) and A-Ci curves revealed similar differences between sun and shade leaves.

## Variation in intercellular and chloroplastic CO~2~ concentrations
Higher rates of g~s~ in shade leaves under low and high light lead to significant increases in C~i~ compared to sun leaves (Figure `r ciccfig`a). The chloroplast CO~2~ partial pressure was comparable between shade leaves when measured at both low and high light conditions (Figure `r ciccfig`c). In sun leaves C~c~ was significantly lower than shade leaves as a likely consequence of a lower C~i~. The drawdown of CO~2~ from intercellular spaces to the chloroplast, C~i~-C~c~, measures the coordination between g~m~ and *A* [@caemmerer2014temperature]. This drawdown was similar between sun and shade leaves measured at their local light environment and increased marginally in shade leaves at high light (Figure `r ciccfig`c). This was the result of the proportional relationship between g~m~ and *A* across all leaves. The CO~2~ drawdown from C~a~ to C~i~ and C~i~ to C~c~ were both relatively stable across the range of temperatures measured and gas exchange campaigns (Figure `r get2fig`c and `r get1fig`c, respectively).

# Discussion
Here we show that *A* in leaves within canopies of *Eucalyptus tereticornis* are limited by their local light environment, however, shade leaves increased rates of leaf C gain exceeding sun leaves when light availability increased. Although shade leaves in lower light environments exhibited relatively high g~s~ and c~i~, it was rapid increases in g~m~ under periods of high light availability that led to this up-regulation of *A*. Although we know shade leaves experience transient periods of sun and shade [@pearcy1990sunflecks], a lack of empirical data within tree canopies currently impedes our ability to predict whole canopy C gain. These findings offer new insights into how aspects of leaf physiology may be optimized differently in sun and shade leaves and reveal how the total leaf CO~2~ conductance pathway should be accounted for when testing optimizations of canopy C uptake in future studies. Additionally, with measurements recorded across a large natural range of air temperatures only minimal effects a +3 &deg;C warming treatment were detected on leaf physiology.

## Resource distribution and photosynthetic capacity
The allocation of N~a~ constrains *A* and is thus a key trait in determining the relative contribution of individual leaves to canopy C gain. Decreasing light availability should decrease the investment into photosynthetic enzyme within a canopy (Mooney and Gulmon 1979). As a result, acclimation of photosynthetic capacity to irradiance is typically reflected in the key photosynthetic biochemical parameters Vc~max~ and J~max~ [@farquhar1980biochemical]. Our data agree with these conventional conclusions as the distribution of N~a~, both measures of photosynthetic capacity and *A* were all reduced in shade leaves. Leaf mass area, however, was not different between sun and shade leaves. This could be due to leaf formation under comparative light conditions or possible differences in total non-structural carbohydrates contents between leaf types.

\
Photosynthesis is also limited by the ability to supply water to the upper canopy. Ultimately, the ability of a trees hydraulic architecture to supply water to foliage across increasing pathlengths affects productivity and survival [@sellin2008effects]. Using a two-leaf model, Peltoniemi et al. [-@peltoniemi2012co] theorizes that optimal N distribution will be proportional to light distribution only if K~l~ is also optimally distributed. In this study, variation in leaf N distribution and *A* rates were not associated with subsequent changes in K~l~ between sun and shade leaves. Thus, no direct relationship between water supply and N distribution or *A* within the canopy were detected.

\
Unexpected higher rates of g~s~ in shade leaves compared to sun leaves led to decreased ITE throughout the experiment. Additionally, consistently higher leaf $\delta$^13^C in shade leaves suggests that this pattern was likely prevalent long term. From a canopy perspective this pattern in water use efficiency initially appears to be detrimental to C gain as *A* in sun leaves was characterized by low rates of g~s~ and low C~i~. Relative to the differences in *A* between leaf types, higher rates of g~s~ in shade leaves appear to exhibit inefficient water use. As whole canopy C gain integrates the efficiency of all leaves, this begs the question of why shade leaves maintained a lower ITE compared to sun leaves.

## Physiological behaviour of sun and shade leaves
The pattern of inefficient water use in shade leaves is important as we hypothesized that g~s~ and *A* would be proportional across sun and shade leaves. In sun leaves, *A* and g~s~ were strongly correlated, exhibiting behaviour broadly consistent with optimal stomatal theory. However, lower rates of *A* in shade leaves were not coupled with decreases in g~s~, explaining the observed decreases in ITE. This is significant as optimal stomatal regulation to balance C gain with water loss has been reported across a wide range of ecosystems and plant functional types; however, empirical data is often collected only on full sun leaves [see @prentice2014balancing; @lin2015optimal]. As a result, the often used economic framework of balancing costs of using water versus N allocation to predict *A* [@wright2003least] may break down when considering all leaves within a tree canopy.

\
It is possible that reducing stomatal response time, by sustaining higher g~s~, is a strategy to take advantage of high light quickly in shade leaves [@tausz2005dynamic]. Evidence from this study supports this hypothesis, as shade leaves increased A equivalent or even outperforming sun leaves under identical light intensity. Transpiration-induced cooling in shade leaves, by keeping stomata open, has also been suggested as an effective strategy to reduce sunfleck induced thermal load [@schymanski2013stomatal]. This is because rapid increases in leaf temperature with sunflecks have been shown to inhibit C gain [@leakey2003high]. However, this response likely occurs at very high air temperatures and may not explain the observed g~s~ in shade leaves across the large natural range of temperatures included in this study. How prevalent each of these strategies are within tree canopies is still unknown, as empirical studies assessing photosynthetic responses to sunflecks generally focus on seedlings [@kuppers1993leaf;@pepin1997rates;@leakey2002relative] and understory plants, often in deep shade [@chazdon1991importance; @allen2000stomatal; @brantley2009contribution]. Thus, our findings highlight a critical need for empirical measurements of shade leaves under dynamic light environments in order to accurately scale C gain from leaf to canopy [see @de1997simple].

\
We found that *A* and g~m~ scaled positively across leaf types and, surprisingly, increased rapidly (within minutes) and proportionately when light intensity was increased in shade leaves.  Research has suggested that aquaporins can facilitate increases in the CO~2~ permeability of the cell membranes resulting in rapid modulation of g~m~ [@hanba2004overexpression; @heinen2009role; @li2014plant]. This provides a potential explanation for the observed rapid increases in g~m~ but the impacts of aquaporins on g~m~ are untested in leaves of tree species. Our findings support a growing wealth of evidence that g~m~ is highly variable and can respond to environmental variables [@flexas2008mesophyll]. Here we provide empirical data showing g~m~ not only varies within a canopy but the up-regulation of g~m~ plays a critical role in the photosynthetic response of shade leaves to sunflecks.

\
If shade leaves "lie in wait" for sunflecks then perhaps we should consider an alternate leaf economic strategy to maximize C gain, beyond conventional trade-offs associated with canopy resource distribution. This is because the role of g~s~ in regulating photosynthetic induction impacts the capacity of a leaf to utilize sunflecks [@way2012sunflecks]. If the valuation of sunflecks as a C resource is large enough then costs of sub-optimal stomatal behaviour could be offset over the leaf lifespan. For example, the potential C gain in leaves where sunflecks constitute a large proportion of total daily PFFD may be large enough to tolerate decreases in ITE However, accounting for the heterogeneous nature of light within a canopy remains a current challenge for empirical and modelling studies. Thus, models which predict leaf photosynthesis from N distribution within a canopy will be incomplete unless inclusion of canopy light extinction and the integration of sunflecks on shade leaves are included [@de1997simple].

## Conclusions
Here we show that dynamic physiological responses of shade leaves to altered light environments has important implications for upscaling leaf level measurements to the canopy. Although resource allocation constrains leaf photosynthetic capacity it is the physiological behaviour of individual leaves which actually determine C gain. These findings suggest that current theories of leaf optimal behaviour should be extended to include dynamic light environments, which will have important implications for process based models that predict canopy C gain from rates of leaf photosynthesis. Furthermore, the dynamic nature of g~m~ cannot be simply parameterized in tree growth models and possibly should be excluded until it can be represented properly. Additional empirical data, across multiple tree species, are needed to determine both the mechanisms and the capacity of g~m~ to rapidly increase CO~2~ drawdown. To improve our ability to predict whole canopy C gain future research should prioritize the incorporation of both sun and shade leaf physiology, which may be optimized differently.

# List of Tables
**Table 1**. *Eucalyptus tereticornis* leaf morphological and physiological traits between full sun and shade leaves under ambient and elevated temperature treatments. Leaf mass per area, N~a~, $\delta$^13^C, $\Psi$~pd~, $\Psi$~l~ and K~l~ values represent treatment mean (&pm; 1 standard error) across measurement campaigns (n=6). Values of Vc~max~ and J~max~ are treatment mean (&pm; 1 standard error) from AC~i~ curves measured in each chamber at saturating light. Units of LMA and Leaf N~area~ are g m^2^, K~l~ is mmol m^-2^ s^-1^ MPa, WP is MPA and $\delta$^13^C is &permil;. Different letters represent significant differences between leaf type and temperature treatments. The P value represents the overall effect between each unique combination of leaf type and temperature treatment for each trait.

\
**Table 2**. Responses of *Eucalyptus tereticornis* leaf gas exchange parameters for sun and shade leaves under ambient and elevated temperature treatments. Each value reflects the mean (&pm; 1 standard error) for each treatment across gas exchange campaigns (n=6). Units for A and E are $\mu$mol m^-2^ s^-1^, for g~s~ and g~m~ are mol m^-2^ s^-1^ and for VPD are kPa. Different letters represent significant differences between leaf type, light environment and temperature treatments. The P value represents the overall effect between each unique combination of leaf type, light environment and temperature treatment for each parameter.

# List of Figures

**Figure 1**. Bars represent the local light environment for sun and shade leaves during six gas exchange campaigns from October 2013 to April 2014. Means &pm; 1 standard error represent integrated PPFD, measured with a ceptometer, at the canopy height of each selected leaf. Each date represents the starting date for each measurement campaign. Points represent the maximum daily air temperature during each campaign period.

\
**Figure 2**. (a) ACi curves for sun and shade leaves at elevated (ET) and ambient (AT) temperature treatments. ACi curves were measured once on all trees, in February 2014, at 25°C and at saturating light (1800 $\mu$mols m^-1^ s^-1^). (b) The relationship between Vc~max~ and mean leaf N~a~ for each chamber, including sun leaves and shade leaves at low light. (c) The relationship between A and leaf N~a~ for sun and shade leaves measured under their ambient light and temperature conditions. For (b,c) the dashed line represents the significant linear model fit for all leaves, with a marginal and conditional R^2^ of 0.28 and 0.35 for (b) and 0.24 and 0.33 for (c).

\
**Figure 3**. The response of A to g~s~ (a) and g~m~ (b) for sun leaves measured at high light and shade leaves measured at both low and high light under their respective elevated and ambient temperature treatments. Lines represent either smoothed regressions from a generalized additive model fit (a) or linear model fits (b). Grey areas are 95% confidence intervals from the mean.

\
**Figure 4**. The mean &pm; 1 standard error of g~s~ (a), g~m~ (b) and A (c) of sun leaves and shade leaves at both low and high light pooled across six measurement dates.

\
**Figure 5**. (a) Response of instantaneous transpiration efficiency (ITE) to VPD for sun leaves and shade leaves at both low and high light with elevated and ambient temperature treatments. (b) The relationship between leaf $\delta$^13^C and leaf N~a~ for sun leaves at high light and shade leaves at low light. For (a) VPD is the leaf to air pressure difference inside the gas exchange cuvette and lines represent predictions from the optimal ITE model with a g~1~ value for each leaf type and treatment. For (b) the dashed line represents the significant linear model fit for all leaves with a marginal and conditional R^2^ of 0.41 and 0.45, respectively.

\
**Figure 6**. Relationship between the observed discrimination of ^13^CO~2~ measured during photosynthesis and measured C~i~/C~a~ for sun leaves measured at high light and shade leaves measured at both low and high light. The solid line represents the theoretical line for C3 plants from Evans et al. [-@evans1986carbon].

\
**Figure 7**. The mean &pm; 1 standard error of intercellular CO~2~ concentration (a), CO~2~ concentration in the chloroplasts (b) and CO~2~ drawdown from substomatal cavities to sites of carboxylation of sun leaves and shade leaves at both low and high light.

\
**Figure S1**. Daily maximum and minimum temperature (a) and total daily PPFD (b) for each chamber across the experiment duration.

\
**Figure S2**. Photosynthetic CO~2~ response (AC~c~) curves for sun and shade leaves at elevated and ambient temperature treatments. C~c~ values were predicted with g~m~, thus curves represent chloroplastic photosynthetic parameters at 25°C and saturating light (1800 $\mu$mols m^-1^ s^-1^).

\
**Figure S3**. Response of A (a), g~m~ (b) and C~i~-C~c~ to leaf temperature for sun leaves and shade leaves at low and high light. Shaded symbols represents each monthly measurement campaign. Solid lines, colored by leaf and light type, are fitted line for the relationship with each parameter and leaf temperature across all measurement campaigns. All parameters with no relationship are fitted with zero slope and the overall mean value for each treatment combination. Weak negative relationships with g~m~ and increasing leaf temperature were detected with sun and shade leaves under their local light environment (R^2^ = 0.16 and 0.08, respectively).

\
**Figure S4**. Response of VPD (a), g~s~ (b) and C~a~-C~i~ to leaf temperature for sun leaves and shade leaves at low and high light. Shaded symbols represents each monthly measurement campaign. Solid lines, colored by leaf and light type, are fitted line for the relationship with each parameter and leaf temperature across all measurement campaigns. All parameters with no relationship are fitted with zero slope and the overall mean value for each treatment combination. Leaf VPD inside the gas exchange cuvette was positively correlated with increasing leaf temperature for sun leaves and shade leaves at low and high light (R^2^ = 0.73, 0.58 and 0.72, respectively)

# Tables {.unnumbered}
**Table 1**. *Eucalyptus tereticornis* leaf morphological and physiological traits between full sun and shade leaves under ambient and elevated temperature treatments. Leaf mass per area, N~a~, $\delta$^13^C, $\Psi$~pd~, $\Psi$~l~ and K~l~ values represent treatment mean (&pm; 1 standard error) across measurement campaigns (n=6). Values of Vc~max~ and J~max~ are treatment mean (&pm; 1 standard error) from AC~i~ curves measured in each chamber at saturating light. Units of LMA and Leaf N~area~ are g m^2^, K~l~ is mmol m^-2^ s^-1^ MPa, WP is MPA and $\delta$^13^C is &permil;. Different letters represent significant differences between leaf type and temperature treatments. The P value represents the overall effect between each unique combination of leaf type and temperature treatment for each trait.

```{r, echo=FALSE}
resourcetab2 <- rbind(resourcetab, c("P value","", as.character(pval)))
greeks=c(alpha='\u03b1', Delta='\u0394', Sigma='\u03A3', mu='\u03bc',delta = '\u03b4')

dust(resourcetab2)%>%
  sprinkle_colnames("Leaf", "Temperature", "LMA", "N~a~", "Vc~max~", "J~max~","K","WP~pre~","WP~mid~", paste0(greeks['delta'],"^13^c")) %>%
  sprinkle(rows= 1, part="head",bold=TRUE)%>%
  sprinkle_print_method("markdown")
```

\
**Table 2**. Responses of *Eucalyptus tereticornis* leaf gas exchange parameters for sun and shade leaves under ambient and elevated temperature treatments. Each value reflects the mean (&pm; 1 standard error) for each treatment across gas exchange campaigns (n=6). Units for A and E are $\mu$mol m^-2^ s^-1^, for g~s~ and g~m~ are mol m^-2^ s^-1^ and for VPD are kPa. Different letters represent significant differences between leaf type, light environment and temperature treatments. The P value represents the overall effect between each unique combination of leaf type, light environment and temperature treatment for each parameter.

```{r, echo=FALSE}
getab2 <- rbind(getab, c("P value","","", as.character(pval2)))
greeks=c(alpha='\u03b1', Delta='\u0394', Sigma='\u03A3', mu='\u03bc',delta = '\u03b4')

dust(getab2)%>%
  sprinkle_colnames("Leaf", "Light","Temperature", "A", "g~s~", "g~m~", "ITE","E","VPD","C~i~", "C~c~") %>%
  sprinkle(rows= 1, part="head",bold=TRUE)%>%
  sprinkle_print_method("markdown")
```

# Figures {.unnumbered}
```{r par, fig.width=6, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
source("master_scripts/par2.R")
```

**Figure 2**.

```{r acicurves, fig.width=8, fig.height=10, warning=FALSE,echo=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
source("master_scripts/aci_vcmaxnitro.R")
```

**Figure 3**.

```{r Agmgs, fig.width=7, fig.height=7, echo=FALSE,warning=FALSE,message=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
source("master_scripts/Agmgs.R")
```

**Figure 4**.

```{r barplots, fig.width=10, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
source("master_scripts/physiology_barplots.R")
```

**Figure 5**.

```{r wateruse, fig.width=7, fig.height=7, echo=FALSE, warning=FALSE, message=FALSE,dev="png", dpi=400}
source("master_scripts/wateruse.R")
```

**Figure 6**.

```{r deltacica, fig.width=6, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
source("master_scripts/delta_cica.R")
```

**Figure 7**.

```{r cicc, fig.width=10, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
source("master_scripts/cicc_bar.R")
```

**Figure 8**.

# Supporting Information {.unnumbered}

```{r airvars, fig.width=7, fig.height=9, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
source("master_scripts/air_vars.R")
```

**Figure S1**

```{r acc, fig.width=6, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
source("master_scripts/ACc_figure.R")
```

**Figure S2**.

```{r gasex_temp3, fig.width=7, fig.height=10, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
source("master_scripts/gasex_temp3.R")
```

**Figure S3**.

```{r gasex_temp4, fig.width=7, fig.height=10, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
source("master_scripts/gasex_temp4.R")
```

**Figure S4**.

# References {.unnumbered}